Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-06 ms.tex
Undoing damage

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}} \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p \tilde{v}_{w}^{2}}{2k} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
{\bf AG missing correction here for black hole velocity dispersion}
where $v_{500} \equiv \tilde{v}_{w}/(500$ km s$^{-1}$).  Hydrostatic equilibrium furthermore determines the value of the stagnation radius (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw^{2}|_{\rs}}\left(\frac{9-\Gamma}{2} \frac{M_{\star}|_{\rs}}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic}
\end{align}
where $n \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm s}}$ is the density power-law slope at $r = r_{\rm s}$.

If the SMBH dominates the gravitational potential at the stagnation radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius resides well inside the sphere of influence ($r_{\rm s} < r_{\rm soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation (\ref{eq:stag_analytic}) simplifies to
\begin{eqnarray}
  &\rs \underset{v_w \gtrsim \sigma}\approx \frac{7}{2}\frac{G \Mbh}{n \tilde{v}_{w}^{2}} \approx \begin{cases}
    12
 \, \pc \,\, \Mbheight v_{500}^{-2}\, \pyear& \text{core} \\
    6
 \, \pc \,\, \Mbheight v_{500}^{-2} \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:stag_simple}
\end{eqnarray}
{\bf AG: missing a minus n here...}
where for the numerical estimate we adopt values of $n \approx 1$ and $n \approx 0.5$ for core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies, respectively, so chosen to match the results of our numerical calculations (Fig.~\ref{fig:stag}).  This expression is similar to that obtained by \citet{Volonteri+11} on more heuristic grounds (their eq.~6).

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}} \tilde{v}_{w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  1.0(\phi/0.1)/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    9.4 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker break radius.  Equation (\ref{eq:conduction}) shows that even in the case of saturated conduction, the neglect of heat conduction results in at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.  Numerical experiments including conductivity confirm this ($\S\ref{sec:numerical}$).

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by other sources of cooling, namely the term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).  Neglecting radiative cooling, this term is balanced at the stagnation radius by the wind heating term, $\dot{q}_{\rm heat} = q \tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius can thus be assessed by comparing the ratio of wind heating to radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   210 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs}) for the gas and stellar densities at the stagnation radius, respectively.  We have approximated the cooling function for $T < 2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6 \text{K}\right)^{-0.7}  $erg cm$^3 $s$^{-1}$ assuming solar metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

To within a constant of order unity, equation (\ref{eq:cooling2}) also equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n kT/2\mu )/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the stagnation radius, as can be derived using the equality 
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \ll t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm
      soi}$ (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a
    function of the ratio of the effective wind heating rate $v_{w}$
    to the stellar velocity dispersion $\sigma$.  Green, orange, and
    blue symbols correspond to different values of $v_{w} =$ 300, 600, and 1200 km s$^{-1}$,
    respectively.  Squares correspond to cusp galaxies ($\Gamma = 0.8$), while
    triangles correspond to cores ($\Gamma = 0.1$). The black curves correspond to the
    analytic prediction from equation (\ref{eq:rs2main}), with solid and dashed curves calculated for parameters $\{\Gamma=0.8, n=1\}$ and $\{\Gamma=0.1,n=0.5\}$, respectively. }
\end{figure}

@@

\begin{eqnarray}
  \left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}}
  &\approx& 12\eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
    Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56} \nonumber \\
  &\approx& 12 M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74}
\label{eq:rIars}
\end{eqnarray}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be
approximated as a steady heating source near the stagnation radius for
extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in
the case of a very young stellar population ($\tau_{\star} \ll t_{\rm
h}$).

@@

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating (e.g., \citealt{Sazonov+04}, \citealt{Ciotti+10}), which provides a volumetric heating rate (\citealt{Gan+14})
\be
\dot{e} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\ee
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$ (e.g., \citealt{Ho99}, \citealt{Eracleous+10}).

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike stellar wind heating, Compton heating depends on radius, scaling as $v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto r^{(\Gamma-n-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.7}$ for core ($\Gamma = 0.8$; n = 1) and cusp ($\Gamma = 0.1$; n = 0.5) galaxies, respectively.  The assumption of our model that the heating parameter be constant with radius is not satisfied, but this radial variation is sufficiently weak that it should not significantly alter our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 130 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 6\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = \dot{M}/\dot{M}_{\rm edd}$ depends on the Eddington ratio as predicted by some models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 2\times 10^{-3} \eta_{0.02}^{0.18}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}

@@

A single burst of star formation does not describe the typical star formation history of most galaxies.  Qualitatively,
smaller galaxies will on average have more recent star formation, resulting in energetic young stellar winds and supernovae which dominate the gas heating budget.  Massive galaxies, on the other hand, will on average possess older stellar populations, with their heating rates dominated by SN Ia (on large radial scales) and SMBH feedback.  We estimate the average value of $\vwO$ as a function of SMBH mass for each heating source by calculating its value using the average cosmic star formation histories of \citet{MosterNaab+:2013a}.  Appendix \ref{app:windheat} describes how the average star formation history is used to determine the stellar wind heating $v_{\rm  w}^{\star}$ and mass return parameter $\eta$ as a function of $M_{\bullet}$ (Fig.~\ref{fig:eta}; top panel).  Star formation histories are also convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating source (stars, MSPs, SNe Ia, and black hole feedback) calculated for the average star formation history of galaxies containing a given black hole mass.  Again, Ia heating is suppressed in low mass galaxies because the Ia radius exceeds the nominal stagnation radius (bottom panel).  MSP heating is again found to be negligible for all black hole masses.  The younger stellar populations characterizing galaxies with $\Mbh\lsim 5\times 10^{8} \Msun$ are dominated by stellar wind and Ia heating, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and $v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most massive galaxies, with $\Mbh\gsim 5\times 10^{8} \Msun$, does the lack of young stellar populations significantly reduce the role of stellar wind heating.   %Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.  
Strikingly, {\it we find that galactic nuclei experiencing the same average star formation history as their host galaxies produce thermally stable flows across all black hole masses}.  {\bf NCS: deleted a sentence referring to old results on thermal instability - it is now correct to say we are thermally stable everywhere, yes?}

>>>>

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}} \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p \tilde{v}_{w}^{2}}{2k} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
{\bf AG missing correction here for black hole velocity dispersion}
where $v_{500} \equiv \tilde{v}_{w}/(500$ km s$^{-1}$).  Hydrostatic equilibrium furthermore determines the value of the stagnation radius (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
  \rs=\frac{G \Mbh}{\densSlope
    v_w^{2}|_{\rs}}\left[\left(\frac{9-\Gamma}{2} -\densSlope\right)
    \frac{\Mstar|_{\rs}}{\Mbh} +\frac{7}{2}-\densSlope\right].
\label{eq:stag_analytic}
\end{align}
where $\densSlope \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm s}}$ is the
density power-law slope at $r = r_{\rm s}$. We find that

\begin{align}
\densSlope \simeq \frac{1}{6} \left(4 \Gamma+3\right).
\label{eq:densSlope}
\end{align}

If the SMBH dominates the gravitational potential at the stagnation
radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius
resides well inside the sphere of influence ($r_{\rm s} < r_{\rm
  soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation
(\ref{eq:stag_analytic}) simplifies to
\begin{eqnarray}
  &\rs \underset{v_w \gtrsim \sigma}\approx \left(\frac{7}{2}-\densSlope\right)\frac{G \Mbh}{\densSlope v_w^2} \approx \begin{cases}
    9
 \, \pc \,\, \Mbheight v_{500}^{-2}\, \pyear& \text{core} \\
    4
 \, \pc \,\, \Mbheight v_{500}^{-2} \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:stag_simple}
\end{eqnarray}
where we have used equation~\eqref{eq:densSlope} to estimate
$\densSlope$: $\densSlope\approx 1$ and $\densSlope \approx 0.6$ for
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies.  This
expression is similar to that obtained by \citet{Volonteri+11} on more
heuristic grounds (their eq.~6).

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(\frac{1}{1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}}\right)
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the
stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius.  Equation (\ref{eq:conduction}) shows that even in the
case of saturated conduction, the neglect of heat conduction results
in at most an order unity correction for expected values of the
saturation parameter $\phi < 0.1$, justifying its neglect in our
analysis.  Numerical experiments including conductivity confirm this
($\S\ref{sec:numerical}$). {\bf AG: If we make this statement then we
  should be sure to include the numerical experiments. The original
  formula was not correct for the saturated case. The numerical values
also changed somewhat. I would appreciate it if someone could double
check this. Also note that there is a weak core/cusp dependence in the
saturated case which I have chosen not to include.}

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by other sources of cooling, namely the term $\propto -q c_{s}^{2} \propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).  Neglecting radiative cooling, this term is balanced at the stagnation radius by the wind heating term, $\dot{q}_{\rm heat} = q \tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius can thus be assessed by comparing the ratio of wind heating to radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   430 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   840 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}

where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$ assuming solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

To within a constant of order unity, equation (\ref{eq:cooling2}) also
equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n
kT/2)/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the
stagnation radius, as can be derived using the equality

\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \ll t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm soi}$
    (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a function of
    the ratio of the effective wind heating rate $v_{w}$ to the
    stellar velocity dispersion $\sigma$.  Green, orange, and blue
    symbols correspond to different values of $v_{w} =$ 300, 600, and
    1200 km s$^{-1}$, respectively.  Squares correspond to cusp
    galaxies ($\Gamma = 0.8$), while triangles correspond to cores
    ($\Gamma = 0.1$). The black curves correspond to the analytic
    prediction from equation (\ref{eq:rs2main}), with solid and dashed
    curves calculated for parameters $\{\Gamma=0.8, \densSlope\simeq1\}$ and
    $\{\Gamma=0.1,\densSlope\simeq0.6\}$, respectively. }
\end{figure}

@@

\begin{align}
  \left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}}&\approx
  \begin{cases}
    8\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56}  & {\rm core}\\
    17\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56} &{\rm cusp}\\
   \end{cases} \nonumber\\
 &\approx 
 \begin{cases}
    8\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74}  & {\rm core}\\
    17\, \eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
     Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74} &{\rm cusp}\\
   \end{cases}
\label{eq:rIars}
\end{align}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be
approximated as a steady heating source near the stagnation radius for
extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in
the case of a very young stellar population ($\tau_{\star} \ll t_{\rm
h}$).

@@

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating (e.g., \citealt{Sazonov+04}, \citealt{Ciotti+10}), which provides a volumetric heating rate (\citealt{Gan+14})
\begin{align}
\dot{e} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\end{align}
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$ (e.g., \citealt{Ho99}, \citealt{Eracleous+10}).

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 27 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 30 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.75}$
for core ($\Gamma = 0.8$; $\densSlope \simeq 1$) and cusp ($\Gamma = 0.1$;
$\densSlope \simeq 0.6$) galaxies, respectively.  The assumption of our model that the heating
parameter be constant with radius is not satisfied, but this radial
variation is sufficiently weak that it should not significantly alter
our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 96 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = 0.1 \dot{M}/\dot{M}_{\rm edd}$ depends on the Eddington ratio as predicted by some models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.4\times 10^{-2} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 6\times 10^{-3} \eta_{0.02}^{0.18}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
{\bf AG: Note that I have included the bolometric correction for now.}

@@

A single burst of star formation does not describe the average star formation histories of most galaxies.  Qualitatively,
smaller galaxies will on average have more recent star formation, resulting in energetic young stellar winds and supernovae which dominate the energy budget.  Massive galaxies, on the other hand, will on average possess older stellar populations, with their heating rates dominated by SN Ia (on large radial scales) and SMBH feedback.  We estimate the average value of $\vwO$ as a function of SMBH mass for each heating source by calculating its value using the average cosmic star formation histories of \citet{MosterNaab+:2013a}.  Appendix \ref{app:windheat} describes how the average star formation history is used to determine the stellar wind heating $v_{\rm  w}^{\star}$ and mass return parameter $\eta$ as a function of $M_{\bullet}$ (Fig.~\ref{fig:eta}; top panel).  Star formation histories are also convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating source (stars, MSPs, SNe Ia, and black hole feedback) calculated for the average star formation history of galaxies containing a given black hole mass.  Again, Ia heating is suppressed in low mass galaxies because the Ia radius exceeds the nominal stagnation radius (bottom panel).  MSP heating is again found to be negligible for all black hole masses.  The younger stellar populations characterizing galaxies with $\Mbh\lsim 5\times 10^{8} \Msun$ results in stellar wind and Ia heating being dominant, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and $v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most massive galaxies $\Mbh\gsim 5\times 10^{8} \Msun$ does the lack of younger stellar populations significantly reduces the role of stellar wind heating.   Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.  Strikingly, {\it we find that galactic nuclei experiencing the same average star formation history as their host galaxies produce thermally stable flows across most black hole masses}.

Realistic variations (``burstiness") in the star formation history will result in lower heating rates.  Figure \ref{app:windheat}  shows the heating rate as a function of average black hole mass also for cases in which the current ($z = 0$) star formation is suppressed by a factor $\epsilon_{\rm burst}$ from its average $z = 0$ value for a characteristic timescale $\tau_{\rm var}$ (for details, see Appendix \ref{fig:}).  For low values of $\tau_{\rm var} \lesssim 10^{7}$ yr we see that $v_{w}^{\star}$ is reduced by at most a factor of $\sim 2$ from its average value, even for very small values of $\epsilon_{\rm burst}=$... , while for $\tau_{\rm var} \gtrsim 10^{8}$ yr, wind heating is reduced significantly.  We conclude that variations in the nuclear star formation rate on timescales short compared to the dynamical timescale of the galaxy are unlikely to significantly affect our results.  {\bf BDM: augment/fix this - its just schematic till nick provides his figure}
